---
title: "Comparative Analysis of Feeding Mode and Microbiome Composition in Poison Frog Tadpoles"
author: "Kayla Weinfurther, Kyle Summers, Adam Stuckert, Ariane L. Peralta"
date: "Last updated on `r format(Sys.time(), '%d %B, %Y')`"
header-includes:
  - \usepackage{array}
  - \usepackage{graphics}
  - \usepackage[utf8]{inputenc}
output: 
  pdf_document:
  fig_caption: true
---

Project Description: Fill out

# Initial Setup
```{r Initial Setup, include=FALSE}
rm(list=ls())
setwd("~/GitHub/TadpoleMicrobiomes/analyses")
se <- function(x, ...){sd(x, na.rm = TRUE)/sqrt(length(na.omit(x)))}
ci <- function(x, ...){1.96 * sd(x,na.rm = TRUE)}

# Code Dependencies
source("../bin/DiversityFunctions.R")
source("../bin/MothurTools.R")
require("vegan")
require("nlme")
require("reshape")
require("ggplot2")
require("ade4")
require("png")
require ("agricolae")
require("tidyr")
require("dplyr")
require("reshape2")
require("picante")
```
#Import Files
## Environmental Data
```{r Import Files - Env, include=FALSE}
# Import Environmental Data
design <- read.csv("../data/poison_tadpole_design.csv", row.names=1)
```
## Bacterial Data
```{r Import Files - Bacteria, include=FALSE}
# Import OTU data
# Import Raw Data
otu.in <- read.otu("../data/tadpole.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.shared")
```

```{r Match design and otu IDs, include=FALSE}
# Correct Sample IDs and Subset File
missing <- setdiff(rownames(design), rownames(otu.in))
design_final <- design[-(which(rownames(design) %in% missing)), ]
dim(design_final)
otu <- otu.in
dim(otu)
otu <- otu[match(rownames(design_final), rownames(otu)),]
all.equal(rownames(otu), rownames(design_final))

# OTU table - remove otus w/ < 2 occurrences across all sites
otu_removal <- otu[, which(colSums(otu) >= 2)]
dim(otu_removal) #went from 8273 to 6114 OTUs

aa <- (rowSums(otu_removal))
aa

# Make Relative Abundance Matrices
dataREL <- otu
for(i in 1:dim(otu)[1]){
  dataREL[i,] <- otu[i,]/sum(otu[i,])
}

tadpole.tax <- read.tax(taxonomy = "../data/tadpole.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.0.03.cons.taxonomy", 
                    format = "rdp", tax.levels = 6, col.tax = 3)
```
# Simple Hypothesis Testing - Microbes
```{r perMANOVA - Bacteria - All Yrs, echo=TRUE}
#PERMANOVA
new.data <-cbind(design_final,dataREL)
adonis = adonis(new.data[,-c(1:3)]~Species*Diet, method = "bray", data = new.data, perm=1000)
adonis
```
# Bacterial Ordinations

## Principal Coordinates Ordination
```{r Ordination (PCoA) - Bacteria - all years, include=FALSE}
# Principal Coordinates Analysis
dataREL.dist <- vegdist(dataREL, method="bray")

pcoa <- cmdscale(dataREL.dist, k=3, eig=TRUE, add=FALSE)
  # Classical (Metric) Multidimensional Scaling; returns PCoA coordinates
  # eig=TRUE returns eigenvalues; k = # of dimensions to calculate

explainvar1b <- round(pcoa$eig[1] / sum(pcoa$eig), 3) * 100
explainvar2b <- round(pcoa$eig[2] / sum(pcoa$eig), 3) * 100
sum.eigb <- sum(explainvar1b, explainvar2b)

explainvar1b #16.0
explainvar2b #13.5
```

```{r Ordination (PCoA) - Bacteria, include=FALSE}
# Principal Coordinates Analysis
#new.data <-cbind(design_final,dataREL)
pcoa.groups <- paste(new.data$Diet, new.data$Species, sep = "_")
pcoa.points <- data.frame(pcoa$points, group = pcoa.groups)

# Calculate Centroids (mean and SE)
pcoa.L.centroids <- melt(pcoa.points, id="group", measure.vars = c("X1", "X2"))
pcoa.centroids <- acast(pcoa.L.centroids, variable ~ group, mean)
pcoa.centroids.se <- acast(pcoa.L.centroids, variable ~ group, se)
pcoa.centroids.sd <- acast(pcoa.L.centroids, variable ~ group, sd)

# Combine
pcoa.cent.dataframe <- cbind(t(pcoa.centroids), t(pcoa.centroids.se))
colnames(pcoa.cent.dataframe) <- c("V1", "V2", "V1e", "V2e")
pcoa.cent.treats <- rownames(pcoa.cent.dataframe)

pcoa.col <- as.factor(sapply(strsplit(pcoa.cent.treats, "_"), `[`, 2)) # Species
pcoa.shape <- as.factor(sapply(strsplit(pcoa.cent.treats, "_"), `[`, 1))  # Diet

#Plot
df1a <- as.data.frame(pcoa.cent.dataframe)
plot1a <- ggplot(df1a, aes(x=V1, y=V2, colour=pcoa.col, shape = pcoa.shape,
                 group = interaction(pcoa.col, pcoa.shape))) + theme_bw() 
plot1a + theme(panel.grid.major = element_blank(), 
               panel.grid.minor = element_blank(), 
               axis.line = element_line(colour = "black")) + 
theme(panel.background = element_blank()) + 
  geom_errorbarh(aes(xmax=V1+V1e, xmin=V1-V1e, height=0.01), colour="black") +    
  geom_errorbar(aes(ymax=V2+V2e, ymin=V2-V2e, width=0.01), colour="black") +
  geom_point(aes(fill=pcoa.col), colour = "black", size=5, stroke = 0.75, show.legend = TRUE) + 
  scale_colour_manual(labels = c("imitator","variabilis"), 
                      values = c("yellow", "darkgreen")) + 
  scale_fill_manual(labels = c("imitator","variabilis"), 
                      values = c("yellow", "darkgreen")) +
  scale_shape_manual(labels = c("detritus","egg fed"),
                     values = c(21, 24)) +
  theme(axis.title = element_text(size=14), axis.text=element_text(size=14),
          axis.text.x = element_text(size=14), 
          panel.border = element_rect(colour = "black", size=1.25)) + 
  theme(axis.ticks.length=unit(0.3,"cm")) + 
  xlab("PCoA 1 (16.0%)") + ylab("PCoA 2 (13.5%)") + 
  labs(fill = "Species", shape = "Diet") +
  guides(fill = guide_legend(override.aes = list(pch=21, size = 4, colour="black")),
         shape = guide_legend(override.aes = list(size = 4, fill="black")))
ggsave("../figures/Ordination_TadpoleMicrobiomes.pdf", plot=last_plot(), device=NULL, path=NULL, scale=1, width=9, height=7, dpi=300, limitsize=TRUE)
```

# Diversity Metrics - Hypothesis Testing
```{r Diversity Metrics - Bacteria, echo=TRUE}
# Rarefy Abundances (min abundance is 13567. We are sampling to 13500)
min(rowSums(otu))
otus.r <- rrarefy(otu, 13500)

# Fisher's Alpha
fisher <- fisher.alpha(otus.r)

# Species Richness
#richness <- rowSums((PWESdata.r >= 1))
richness <- rowSums((otu >= 1))

# Shannon Diversity
shannon <- diversity(otus.r, "shannon")

# Simpson's Evenness
simp.even <- apply(otus.r, 1, simp_even)

#Pielou’s evenness
J <- shannon/log(specnumber(otus.r[,-c(1:1)]))

#combined richness, diversity, evenness
diversity <- cbind(design_final,richness,shannon,simp.even,J)
write.csv(diversity,"../data/diversity.bact.raw.csv")
```

# Diversity Metrics - Hypothesis Testing
```{r Hypothesis Testing - Bacteria, echo=TRUE}
#summary table for bacterial diversity
summary <- diversity %>% group_by(Species, Diet) %>% summarise(mean.richness=mean(richness), se.richness=se(richness), mean.shannon=mean(shannon), se.shannon=se(shannon))
print(summary)
write.csv(summary,"../data/diversity.bact.summary.csv")

library(lmerTest)

richness.lm <- glm(richness ~ Species*Diet, data = diversity)
plot(richness.lm)
richness.lm
anova(richness.lm)

evenness.lm <- glm(simp.even ~ Species*Diet, data = diversity)
plot(evenness.lm)
evenness.lm
anova(evenness.lm)

shannon.lm <- glm(shannon ~ Species*Diet, data = diversity)
plot(shannon.lm)
shannon.lm
anova(shannon.lm)
```

#Plot shannon diversity 
```{r Plot - Shannon Diversity, echo=TRUE}
# Graphing Shannon Diversity
p <- ggplot(diversity, aes(x=Diet, y=shannon, color=as.factor(Species)))+ scale_color_manual(name="Species", values=c("yellow", "darkgreen"), labels = c("imitator", "variabilis")) + stat_summary(fun.data=mean_cl_boot,size=0.75) 
p1=p+geom_smooth(method="lm")
p1 + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "Diet", y = "Shannon Diversity Index (H')") + 
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1)) +
    scale_x_discrete(breaks=c("detritus", "egg fed"))
  
ggsave("../figures/shannon.pdf", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```

#Plot Evenness 
```{r Plot - Simpson's Evenness, echo=TRUE}
# Graphing Simpson's Evenness
p <- ggplot(diversity, aes(x=Diet, y=simp.even, color=as.factor(Species)))+ scale_color_manual(name="Species", values=c("yellow", "darkgreen"), labels = c("imitator", "variabilis")) + stat_summary(fun.data=mean_cl_boot,size=0.75) 
p1=p+geom_smooth(method="lm")
p1 + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "Diet", y = "Simpson's Evenness") + 
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1)) +
    scale_x_discrete(breaks=c("detritus", "egg fed"))
  
ggsave("../figures/simp.even.pdf", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```
#Plot Richness
```{r Plot - OTU richness, echo=TRUE}
# Graphing Richness
p <- ggplot(diversity, aes(x=Diet, y=richness, color=as.factor(Species)))+ scale_color_manual(name="Species", values=c("yellow", "darkgreen"), labels = c("imitator", "variabilis")) + stat_summary(fun.data=mean_cl_boot,size=0.75) 
p1=p+geom_smooth(method="lm")
p1 + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "Diet", y = "OTU Richness") + 
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1)) +
    scale_x_discrete(breaks=c("detritus", "egg fed"))
  
ggsave("../figures/richness.pdf", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```

# Bacterial community indicator species analysis - Diet
```{r Bacteria Indicator Species, include=FALSE}
library("labdsv")

dataREL <- dataREL[, colSums(dataREL) > 0.05]
design.type <- new.data$Diet
bac.ind <- indval(dataREL, design.type)
levels(design.type)
summary(bac.ind)

inds <- which(bac.ind$pval <= 0.05)
bac.indicators <- as.data.frame(matrix(NA, nrow = length(inds), ncol = 4))
colnames(bac.indicators) <- c("OTU", "Cluster", "IndVal", "Prob")

bac.indicators$OTU <- names(inds)
bac.indicators$Cluster <- bac.ind$maxcls[inds]
bac.indicators$IndVal <- bac.ind$indcls[inds]
bac.indicators$Prob <- bac.ind$pval[inds]

ind.tax <- tadpole.tax[which(as.character(tadpole.tax$OTU) %in% bac.indicators$OTU), ]
ind.tax <- ind.tax[match(ind.tax$OTU, bac.indicators$OTU), ]

indicator.bac <- cbind(bac.indicators, ind.tax[, -c(1)])

indicator.bac <- indicator.bac[order(as.numeric(indicator.bac$Cluster)), ]

table(indicator.bac$Cluster)
table(indicator.bac$Phylum)
table(indicator.bac$Cluster)
levels(design.type)

# Export Bacteria Indicator Table
write.table(indicator.bac, "../data/BacterialIndicators_Diet.txt",
            sep="\t", row.names = F, quote = F)
```

# Bacterial community indicator species analysis - Species
```{r Bacteria Indicator Species, include=FALSE}
library("labdsv")

dataREL <- dataREL[, colSums(dataREL) > 0.05]
design.type <- new.data$Species
bac.ind <- indval(dataREL, design.type)
levels(design.type)
summary(bac.ind)

inds <- which(bac.ind$pval <= 0.05)
bac.indicators <- as.data.frame(matrix(NA, nrow = length(inds), ncol = 4))
colnames(bac.indicators) <- c("OTU", "Cluster", "IndVal", "Prob")

bac.indicators$OTU <- names(inds)
bac.indicators$Cluster <- bac.ind$maxcls[inds]
bac.indicators$IndVal <- bac.ind$indcls[inds]
bac.indicators$Prob <- bac.ind$pval[inds]

ind.tax <- tadpole.tax[which(as.character(tadpole.tax$OTU) %in% bac.indicators$OTU), ]
ind.tax <- ind.tax[match(ind.tax$OTU, bac.indicators$OTU), ]

indicator.bac <- cbind(bac.indicators, ind.tax[, -c(1)])

indicator.bac <- indicator.bac[order(as.numeric(indicator.bac$Cluster)), ]

table(indicator.bac$Cluster)
table(indicator.bac$Phylum)
table(indicator.bac$Cluster)
levels(design.type)

# Export Bacteria Indicator Table
write.table(indicator.bac, "../data/BacterialIndicators_Species.txt",
            sep="\t", row.names = F, quote = F)
```